---
title: 'Trial HPLC SN-38/G from methanol standards and LB extractions'
output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged

---

******************************************************* 

#### Authored: Kathy N. Lam 

#### Created: September 10, 2020 

#### Updated: `r format(Sys.Date(), "%B %d, %Y")`

### Set up

<style type="text/css">
div.main-container {
  max-width: 100% !important;
}
.tocify {
  max-width: 100% !important;
}
.toc-content {
  padding-left: 0px !important;
}
</style>

```{r}
library(purrr) #for reading files into tibble
library(gtools) #for natural sorting
library(stringr) #for regex
library(cowplot) #for multi panel plots
library(tidyverse) #for data wrangling and plotting
```

### Read sample data

```{r}
samples = readr::read_csv("samples.csv") 
print(samples, n=12)
```

### Read and wrangle signal files

```{r}
#collect data for all files
path = "data"
(files = list.files(path, pattern = ".*SIGNAL.*.CSV"))
(signal_data = tibble(Filename = files) %>% 
  mutate(File_contents = purrr::map(Filename, function(x) read.csv(file.path(path, x), header = F, sep="\t", fileEncoding = "UTF-16"))) %>% #readr cannot handle UTF16?
  unnest(cols=c(File_contents)) %>%
  rename(Time=V1, Absorbance=V2) %>%
  mutate(SampleNumber = stringr::str_match(Filename, ".*SAMPLE([0-9]{1,2}).*")[,2]) %>% #second item is capture group; first is whole string match
  select(Filename, SampleNumber, Time, Absorbance) %>%
  mutate(SampleNumber=as.numeric(SampleNumber)) %>%
  left_join(samples, by="SampleNumber") %>%
  mutate(SampleNumber=factor(SampleNumber, levels=gtools::mixedsort(unique(SampleNumber))))
)
```

### Read and wrangle integration files

```{r}
#collect data for all files
path = "data"
(files = list.files(path, pattern = ".*INTEGRATION.*.CSV"))
(integration_data = tibble(Filename = files) %>% 
  mutate(File_contents = purrr::map(Filename, function(x) read.csv(file.path(path, x), header = F, sep="\t", fileEncoding = "UTF-16"))) %>% #readr cannot handle UTF16?
  unnest(cols=c(File_contents)) %>%
  rename(Peak=V1, Time=V2, Area=V3, Height=V4) %>%
  mutate(SampleNumber = stringr::str_match(Filename, ".*SAMPLE([0-9]{1,2}).*")[,2]) %>% #second item is capture group; first is whole string match
  select(Filename, SampleNumber, Peak, Time, Area, Height)  %>%
  mutate(SampleNumber=as.numeric(SampleNumber)) %>%
  left_join(samples, by="SampleNumber") %>%
  mutate(SampleNumber=factor(SampleNumber, levels=gtools::mixedsort(unique(SampleNumber))))
)
```

### Plot signal and integration 

```{r}
#make dir for figures
suppressWarnings(dir.create("figures"))
```


```{r, fig.width=14, fig.height=5.5}
sample_subset = samples %>%
    filter(SampleNumber %in% signal_data$SampleNumber) %>%
    mutate(SampleNumber=factor(SampleNumber, levels=mixedsort(unique(SampleNumber))))

ggplot() +
  geom_line(data=signal_data, aes(x=Time, y=Absorbance, colour=Compound), size=1, alpha=1) +
  scale_colour_manual(values = c("indianred", "steelblue", "black")) +
  geom_text(data=sample_subset, aes(x=12.5, y=1000, label=paste0(Solvent, "\n", Compound, "\n", Concentration, "\n", PercentDMSO, "% DMSO")), size=3, hjust=1) +
  geom_text(data=integration_data, aes(x=Time, y=Height+300, label=paste0(round(Time,2)," (",round(Area,0), ")")), size=2.5, angle=90) +
  scale_y_continuous(limits=c(-300, 1400), name="Absorbance (360 nm)") +
  scale_x_continuous(breaks=seq(0,13,2), name="Time (min)") +
  facet_wrap(~SampleNumber, ncol=2) +
  theme_linedraw(14) +
  facet_wrap(~SampleNumber, ncol=5) +
  theme(panel.grid.minor=element_blank())

ggsave("figures/2020-09-09_HPLC_all_samples_uncurated.png")
ggsave("figures/2020-09-09_HPLC_all_samples_uncurated.pdf")
```

### Plot signal and integration - curated 

```{r, fig.width=14, fig.height=5.5}
sample_subset = samples %>%
    filter(SampleNumber %in% signal_data$SampleNumber) %>%
    mutate(SampleNumber=factor(SampleNumber, levels=mixedsort(unique(SampleNumber))))

#manually curate irrelevant peaks per sample identified by HPLC software
integration_curated = integration_data %>%
    filter(SampleNumber %in% sample_subset$SampleNumber) %>%
    filter(Area > 100) %>%
    filter(Time > 3 & Time < 6) %>%
    filter(SampleNumber != 5 | Time > 5) %>% 
    filter(SampleNumber != 6 | Time > 4) %>%
    filter(SampleNumber != 7) %>%
    filter(SampleNumber != 8 | Time < 5.2) %>%
    filter(SampleNumber != 9 | (Time > 4 & Time < 4.1)) %>%
    filter(SampleNumber != 11 | (Time > 5 & Time < 5.2)) %>%
    filter(SampleNumber != 12 | (Time < 4.1)) 

ggplot() +
  geom_line(data=signal_data, aes(x=Time, y=Absorbance, colour=Compound), size=1, alpha=1) +
  scale_colour_manual(values = c("indianred", "steelblue", "black")) +
  geom_text(data=sample_subset, aes(x=12.5, y=1000, label=paste0(Solvent, "\n", Compound, "\n", Concentration, "\n", PercentDMSO, "% DMSO")), size=3, hjust=1) +
  geom_text(data=integration_curated, aes(x=Time, y=Height+300, label=paste0(round(Time,2)," (",round(Area,0), ")")), size=2.5, angle=90) +
  scale_y_continuous(limits=c(-300, 1400), name="Absorbance (360 nm)") +
  scale_x_continuous(breaks=seq(0,13,2), name="Time (min)") +
  facet_wrap(~SampleNumber, ncol=2) +
  theme_linedraw(14) +
  facet_wrap(~SampleNumber, ncol=5) +
  theme(panel.grid.minor=element_blank())

ggsave("figures/2020-09-09_HPLC_all_samples.png")
ggsave("figures/2020-09-09_HPLC_all_samples.pdf")
```

### Plot methanol high conc.

```{r, fig.width=14, fig.height=3}
MeOH100uM = signal_data %>%
  filter(SampleNumber=="5" | SampleNumber =="6") %>%
  filter(Time<10)

integrationsubset = integration_curated %>%
    filter(SampleNumber %in% MeOH100uM$SampleNumber) 

(p1=ggplot() +
  geom_line(data=MeOH100uM, aes(x=Time, y=Absorbance, colour=Compound), size=1, alpha=0.75) +
  scale_colour_manual(values = c("indianred", "steelblue")) +
  geom_text(data=integrationsubset, aes(x=Time, y=Height, label=paste0(round(Time,2)," (",round(Area,0), ")")), size=3) +
  scale_y_continuous(limits=c(-200, 700), name="Absorbance (360 nm)") +
  scale_x_continuous(breaks=seq(0,10), name="Time (min)") +
  theme_linedraw(14) +
  theme(panel.grid.minor = element_blank(), plot.title=element_text(size=14)) +
  ggtitle("Standards prepared in methanol (100 uM)")
)

ggsave("figures/2020-09-09_HPLC_standards_methanol_100uM.png")
ggsave("figures/2020-09-09_HPLC_standards_methanol_100uM.pdf")
```

### Plot LB low conc.

```{r, fig.width=14, fig.height=3}
LB10uM = signal_data %>%
  filter(SampleNumber =="7" | SampleNumber =="8" | SampleNumber =="9") %>%
  filter(Time<10)

integrationsubset = integration_curated %>%
    filter(SampleNumber %in% LB10uM$SampleNumber)

(p2=ggplot() +
  geom_line(data=LB10uM, aes(x=Time, y=Absorbance, colour=Compound), size=1, alpha=0.65) +
  scale_colour_manual(values = c("indianred", "steelblue", "black")) +
  geom_text(data=integrationsubset, aes(x=Time, y=Height, label=paste0(round(Time,2)," (",round(Area,0), ")")), size=3) +
  scale_y_continuous(limits=c(-200, 200), name="Absorbance (360 nm)") +
  scale_x_continuous(breaks=seq(0,10), name="Time (min)") +
  theme_linedraw(14) +
  theme(panel.grid.minor = element_blank(), plot.title=element_text(size=14)) +
  ggtitle("Spike-in extractions from LB (10 uM and two-fold concentrated)")
)

ggsave("figures/2020-09-09_HPLC_LB_10uM.png")
ggsave("figures/2020-09-09_HPLC_LB_10uM.pdf")
```

### Combine methanol and LB

```{r, fig.width=14, fig.height=6}
cowplot::plot_grid(p1, p2, align="v", nrow=2)
ggsave("figures/2020-09-09_HPLC_methanol_LB_combined.png")
ggsave("figures/2020-09-09_HPLC_methanol_LB_combined.pdf")
```

### Notes

* SN-38 may have been prepared at lower concentration than SN-38G, judging from the standards (about 2-fold difference)
* However, extraction still is poorer from LB of SN-38
* Repeat with replicates; also see notes in spreadsheet for what to try next
* Also, Moriah says something is up with the instrument or column; signal should not be dropping as it is, and other wavelengths have strange high signals


### Log session

```{r}
R.version
sessionInfo()
```

































